Thermodynamic fragments in magic state theories

### 2021-02-18 magic_paper.tex
Checked Sec.III and IV. Comments left where I am unsure about what to write

@@

The free operations should send the set of free states $\Fmax$ into itself and completely preserve the non-negativity of the states, in the sense that $\E \in \Omax$ iff $(\id_d \otimes \E ) \sigma \in \stab$ for all odd prime dimensions $d$ whenever $\sigma \in \Fmax$.
It is shown by Wang \textit{et al.}~\cite{cit:wang} that $\Omax$ coincides with the set of operations $\E$ that correspond to stochastic Wigner distributions, 
\begin{equation}
    \Omax = \{ \E: \W[\bmy|\bmx]{\E} \geq 0 \text{ for all } \bmx, \bmy \in \cal{P}_d\}.
\end{equation}

Any magic theory $\R = (\F, \O)$ is a subtheory of $\Rmax$ as explained in~\cref{sec:intro}, and as such it falls under this new stochasticity setting. 
For technical simplicity in what follows we assume that $\F$ is a closed set, and note that $\F_{\rm{ max}}$ is itself a closed set, since it is specified by a finite set of linear constraints of the form $\tr[ L\rho] \geq 0$ with operators $L \in \cal{B}(\cal{H})$ ensuring that the state is positive and Wigner-positive.

Given this context we now define the following key notion, that is central to our analysis.
\begin{definition}[\textbf{$\boldsymbol\sigma$-fragment}]\label{def:sigmafrag}
   Given a resource theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the resource theory $\R_\sigma = (\F, \O_\sigma)$, where the free operations are restricted to the ones that leave $\sigma$ invariant, namely
    \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \}.
    \end{equation}
\end{definition}

With this basic notion defined, we now show that any resource theory of magic can be faithfully subdivided into $\sigma$--fragments, in such a way that any problem of interconversion in the parent magic theory $\R$ can be analysed across the different fragments.

@@

    The state interconversion result follows immediately.
\end{proof}

The zoo of all magic operation classes is summarised in ~\cref{fig:zoo}.
Completely positive-Wigner-preserving operations~\cite{cit:wang} form the operation class $\Omax$.
Therefore, $\sigma$--fragments cover this theory of magic exactly and any magic subtheory is contained within this cover.
In particular, the stabilizer operations $\so$ are contained within $\Omax$.
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.47
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) involve operations denoted by the two yellow regions, with every other established magic theory  between them.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment $\bmd$-majorization can be used allowing for an intricate, but tractable approach towards the study of magic state interconversion.
    }
    \label{fig:zoo}
\end{figure}

@@

Majorization is a collection of powerful tools that has recently found many applications in quantum information theory \nick{CITE}.
It can describes the \nick{disorder / non-uniformity} of distributions that undergo stochastic transformations.

To formally state majorization results, we first denote by $\stoch$ the set of $(d \times d)$ stochastic matrices that preserve the probability vector $\bmd$. \nick{Should we introduce notation directly in the magic setting?}
Specifically, for any $S \in \stoch$, all matrix elements are non-negative, all rows sum to $1$ and $S\bmd = \bmd$.
\iffalse
\begin{enumerate}
    \item $S_{ij} \geq 0$ for all $i, j \in \mathbb{Z}_d$;
    \item $\sum\limits_{j=1}^n S_{ij} = 1$ for all $i \in \mathbb{Z}_d$;
    \item $S\bmd = \bmd$.
\end{enumerate}
\fi
The set $\stoch$ forms a group under matrix multiplication for all $\bmd$ with positive components.

Majorization finds an important application on quantum thermodynamics in the absence of coherence.
The use of majorization in this setting provides useful intuition for our purposes.
At any given temperature $\beta^{-1}$, the thermal state $\gamma_\beta$ is thermodynamically the most disordered state. 
Thermal operations are defined as operations that cannot extract energy from the Gibbs state, $\E(\gamma_\beta) = \gamma_\beta$.
Convertibility between states via thermal operations is equivalent to a stochasticity condition on the energy level populations of the states \nick{CITE}.
Roughly, the statement is that there exists a thermal operation $\E$ such that $\tau = \E(\rho)$ if and only if there exists a a matrix $S \in \stoch$ such that $\bm{q} = S\bm{p}$, where $\bm{q}, \bm{p}$ and $\bmd$ and the energy level population vectors of $\tau, \rho, \gamma_\beta$ respectively.

Drawing intuition from this setting, we can define majorization as follows.
\begin{definition}[\textbf{$\boldsymbol\bmd$-majorization}]\label{def:dmajor}
    Given $\bmx, \bmy, \bmd \in \reals^d$, such that the components of $\bmd$ are positive, $\bmy$ is said to $\bmd$-majorize $\bmx$, iff there exists a matrix $S \in \stoch$ such that $\bmx = S\bmy$.
\end{definition}
We denote this pre-order by $\bmx \prec_{\bmd} \bmy$.
If $\bmd = \frac{1}{d}\bm{1}$, the $d$-dimensional uniform distribution, then $\stoch$ is the set of doubly stochastic matrices and we retrieve the familiar notion of majorization in entanglement theory. \nick{CITE}

The pre-order $\prec_{\R_\sigma}$ of the $\sigma$--fragment $\R_\sigma = (\F, \O_\sigma)$ between $d$-dimensional states corresponds to the majorization pre-order $\prec_{\W{\sigma}}$ between their $d^2$-dimensional Wigner distributions.
For simplicity we shall merge the notation into $\prec_\sigma$, as there is little risk of confusion.

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible \nick{only if}  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
\end{theorem}
\begin{proof}
Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
Since $\sigma$ is \nick{full-rank} and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
However the majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- this enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

This approach can also provide \emph{lower bounds} on distillation, however now more structure about the specific free operations must be included. 
\nick{This is more involved and we discuss this later in the paper.
Also mention relative majorization somewhere}

\subsection{Majorization computations in $\sigma$--fragments}
A numerically efficient reformulation of $\bmd$-majorization is provided by Lorenz curves.
Let the vector $\bmu^\downarrow$ denote the vector $\bmu \in \reals^d$ with its components arranged in non-increasing order.
\begin{definition}[\textbf{Lorenz curve}]\label{def:lc}
    Let $\bmz, \bmd \in \reals^d$, where the components of $\bmd$ are positive with $D = \sum_{i=1}^d d_i$ and denote by $\tilde{\bmz} \coloneqq (z_1/d_1, \dots, z_d/d_d)^T$ the vector of ratios between the corresponding components of $\bmz$ and $\bmd$.

@@

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmz}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmz}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k z_{\pi(i)} \right) \in \mathbb{R}^2.
    \end{equation}
    We define the set of points on this curve, $\lc{\bmz}{\bmd}(x),\ x \in [0,1]$, as the \emph{Lorenz curve} of vector $\bmz$ with respect to $\bmd$.
\end{definition}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmz}{\bmd}$ and $\lc{\bmz \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmz}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\tilde{\bmz}^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.
Finally, points on the interior of the Lorenz curve connecting line segments of different slopes are called \emph{elbows}.

A vector $\bmy$ \emph{$\bmd$--majorizes} another vector $\bmx$ if and only if the Lorenz curve $\lc{\bmy}{\bmd}$ lies above Lorenz curve $\lc{\bmx}{\bmd}$, thus reducing $\bmd$--majorization into a finite set of inequalities.
\begin{theorem}\label{thm:dmajor}
    Let $\bmx, \bmy, \bmd \in \reals^d$, such that the components of $\bmd$ are positive. 
    Then, $\bmx \prec_{\bmd} \bmy$ if and only if $\lc{\bmx}{\bmd}(x) \leq \lc{\bmy}{\bmd}(x)$ for all $x \in [0,1]$ with strict equality at $x=1$.
\end{theorem}
A restatement of the theorem including more equivalent conditions and a proof are provided in~\cref{app:major}.
An example of comparison between different Lorenz curves is illustrated in~\cref{fig:lctoy}.
\begin{figure}
    \centering
    \includegraphics[height=5cm]{figs/lctoy.pdf}
    \caption{Example of different Lorenz curves for quasi-probability vectors under $\bmd$-majorization.
    Vectors $\bmy$ and $\bmd$ are simply probability distributions.
    The curve corresponding to vector $\bmd$ is always the line segment connecting $(0,0)$ and $(1,1)$, so that any other Lorenz curve lies above it, for example $\bmx \prec_{\bmd} \bmd$.
    Curves $L_k(\bmx|\bmd)$ and $L_k(\bmy|\bmd)$ intersect, so $\bmx \nprec_{\bmd} \bmy$ as well as $\bmy \nprec_{\bmd} \bmx$.
    \nick{Recast in terms of magic - replace with~\cref{fig:test}?}
    }
    \label{fig:lctoy}
\end{figure}

>>>>

\nick{Should I define things in (a) 1-copy case only (b) $n$-copy case only (c) 1-copy case, then $n$-copy case ?}

@@

The free operations should send the set of free states $\Fmax$ into itself and completely preserve the non-negativity of the states, in the sense that $\E \in \Omax$ iff $(\id_d \otimes \E ) \sigma \in \stab$ for all odd prime dimensions $d$ whenever $\sigma \in \Fmax$.
It is shown by Wang \textit{et al.}~\cite{cit:wang} that $\Omax$ coincides with the set of operations $\E$ with stochastic Wigner distributions, 
\begin{equation}
    \Omax = \{ \E: \W[\bmy|\bmx]{\E} \geq 0 \text{ for all } \bmx, \bmy \in \cal{P}_d\}.
\end{equation}

Any magic theory $\R = (\F, \O)$ that has been proposed is a subtheory of $\Rmax$ in the sense that $\F \subseteq \Fmax$ and $\O \subseteq \Omax$, and as such it falls under this new stochasticity setting. 
For technical simplicity in what follows we assume that $\F$ is a closed set, and note that $\F_{\rm{ max}}$ is itself a closed set, since it is specified by a finite set of linear constraints of the form $\tr[ L\rho] \geq 0$ with operators $L \in \cal{B}(\cal{H})$ ensuring that the state is positive and Wigner-positive.

Given this context we now define the following key notion, that is central to our analysis.
\begin{definition}[\textbf{$\boldsymbol\sigma$-fragment}]\label{def:sigmafrag}
   Given a resource theory of magic $\R = (\F, \O)$, the \emph{$\sigma$--fragment of $\R$} is the resource theory $\R_\sigma = (\F_\sigma, \O_\sigma)$, where $\F_\sigma = \{\sigma\}$ and the free operations are restricted to the ones that leave $\sigma$ invariant,
    \begin{equation}
        \O_\sigma \coloneqq \{ \E \in \O: \E(\sigma) = \sigma \}.
    \end{equation}
\end{definition}

Note that $\F = \bigcup_{\sigma \in \F} \F_\sigma$ trivially.
With this basic notion defined, we now show that this union holds for free operations as well and therefore that any resource theory of magic can be faithfully subdivided into $\sigma$--fragments, in such a way that any problem of interconversion in the parent magic theory $\R$ can be analysed across the different fragments.

@@

    The state interconversion result follows immediately.
\end{proof}
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.47
        ]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ intro $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) involve operations denoted by the two yellow regions, with every other established magic theory  between them.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment $\bmd$-majorization can be used allowing for an intricate, but tractable approach towards the study of magic state interconversion.
    }
    \label{fig:zoo}
\end{figure}

The zoo of all magic operation classes is summarised in ~\cref{fig:zoo}.
Completely positive-Wigner-preserving operations~\cite{cit:wang} form the maximal operation class $\Omax$.
This theory of magic is thus covered by $\sigma$--fragments exactly and every other magic theory is contained within this cover.
In particular, the stabilizer operations $\so$ are contained within $\Omax$.

@@

Majorization is a collection of powerful tools that has recently found many applications in quantum information theory \nick{CITE}.
It describes the \nick{disorder / non-uniformity} of distributions that undergo stochastic transformations.

To formally state majorization results, we first denote by $\stoch$ the set of $(d \times d)$ stochastic matrices that preserve the probability vector $\bmd$.
Specifically, for any $S \in \stoch$, all matrix elements are non-negative, all rows sum to $1$ and $S\bmd = \bmd$.
\iffalse
\begin{enumerate}
    \item $S_{ij} \geq 0$ for all $i, j \in \mathbb{Z}_d$;
    \item $\sum\limits_{j=1}^n S_{ij} = 1$ for all $i \in \mathbb{Z}_d$;
    \item $S\bmd = \bmd$.
\end{enumerate}
\fi
The set $\stoch$ forms a group under matrix multiplication for all $\bmd$ with positive components.

Majorization exactly describes quantum thermodynamics in the absence of coherence.
The use of majorization in this setting provides vital intuition for our purposes.
At any given temperature $\beta^{-1}$, the thermal state $\gamma_\beta$ is thermodynamically the most disordered state. 
Thermal operations are defined as operations that cannot extract energy from the Gibbs state, $\E(\gamma_\beta) = \gamma_\beta$.
Convertibility between states via thermal operations is equivalent to a stochasticity condition on the energy level populations of the states \nick{CITE}.
Informally, there exists a thermal operation $\E$ such that $\tau = \E(\rho)$ if and only if there exists a matrix $S \in \stoch$ such that $\bm{q} = S\bm{p}$, where $\bm{q}, \bm{p}$ and $\bmd$ and the energy level population vectors of $\tau, \rho, \gamma_\beta$ respectively.

Based on this setting, we define majorization as follows.
\begin{definition}[\textbf{$\boldsymbol\bmd$-majorization}]\label{def:dmajor}
    Given $\bmx, \bmy, \bmd \in \reals^d$, such that the components of $\bmd$ are positive, $\bmy$ is said to $\bmd$-majorize $\bmx$, iff there exists a matrix $S \in \stoch$ such that $\bmx = S\bmy$.
\end{definition}
We denote this pre-order by $\bmx \prec_{\bmd} \bmy$.
If $\bmd = \frac{1}{d}\bm{1}$, the $d$-dimensional uniform distribution, then $\stoch$ is the set of doubly stochastic matrices and we retrieve the familiar notion of majorization from entanglement theory. \nick{CITE}

In any magic theory, we can pick a certain free state, for example a stabilizer and view it as the analog of some thermal state $\gamma_\beta$.
Without loss of generality we can always write $\sigma = \frac{1}{\Z} e^{-\beta H}$ for some $\beta \ge 0$ and Hamiltonian $H$ (either effective or actual)\footnote{Technicalities arise for the case where $\sigma$ is not full rank, but this can be still described via $ \beta \rightarrow 0$ limiting process.}.
It then defines a set of free operations $\O_\sigma$ analogous to the thermal operations that preserve $\gamma_\beta$ the pre-order $\prec_{\R_\sigma}$ between these operations follows the rules of $\bmd$-majorization.

In particular, the pre-order $\prec_{\R_\sigma}$ of the $\sigma$--fragment $\R_\sigma = (\F, \O_\sigma)$ between $d$-dimensional states corresponds to the majorization pre-order $\prec_{\W{\sigma}}$ between their $d^2$-dimensional Wigner distributions.
For simplicity we shall merge the notation into $\prec_\sigma$, as there is little risk of confusion.

\begin{theorem}\label{thm:sigmamajor}
    Let $\R = (\F, \O)$ be a theory of magic. Then $\rho \longrightarrow \tau$ is possible \nick{only if}  $\W{\tau} \prec_\sigma \W{\rho}$ within a $\sigma$--fragment.
\end{theorem}
\begin{proof}
Suppose we can convert $\rho$ into $\tau$ in the magic theory. 
Thus there is some $\O_\sigma$ and some $\E \in \O_\sigma$ such that $\E(\rho) = \tau$, and $\E(\sigma)=\sigma$. 
Therefore, the Wigner distribution of this free operation satisfies $\W{\E} \in \stochw$ and $\W{\E}\W{\rho} = \W{\tau}$. 
Since $\sigma$ is \nick{full-rank} and free, its Wigner distribution is strictly positive in all components, so it directly follows from~\cref{def:dmajor} that $\W{\tau} \prec_{\sigma} \W{\rho}$.
\end{proof}
This result can be understood as an extension of the idea of a magic monotone, where we replace $\M(\tau) \leq \M(\rho)$ with $\W{\tau} \prec_\sigma \W{\rho}$. 
The physical difference between the two expressions is that the majorization ordering occurs in a specific $\sigma$-fragment.
Therefore, majorization constraints can be used to place upper bounds on magic state distillation in a way that allows one to incorporate the physics of the allowed operations -- specifically, it enables us to bound how much magic can be distilled via quantum operations that, for example, preserve the equilibrium state of the system, or via operations that are symmetric about the $Z$-axis of the Bloch sphere.

This approach can also provide \emph{lower bounds} on distillation, however now more structure about the specific free operations must be included. 
\nick{This is more involved and we discuss this later in the paper.}

\nick{Also mention relative majorization somewhere}

\subsection{Majorization computations in $\sigma$--fragments}
The pre-order imposed by $\bmd$-majorization admits a numerically efficient reformulation in terms of Lorenz curves.
Let the vector $\bmu^\downarrow$ denote the vector $\bmu \in \reals^d$ with its components arranged in non-increasing order.
\begin{definition}[\textbf{Lorenz curve}]\label{def:lc}
    Let $\bmz, \bmd \in \reals^d$, where the components of $\bmd$ are positive with $D = \sum_{i=1}^d d_i$ and denote by $\tilde{\bmz} \coloneqq (z_1/d_1, \dots, z_d/d_d)^T$ the vector of ratios between the corresponding components of $\bmz$ and $\bmd$.

@@

    Consider the piecewise linear curve obtained by joining the points $\{(0,0)\} \cup \{ (x_k, \lc{\bmz}{\bmd}(k)) \}_{k=1,\dots,d}$, where
    \begin{equation}\label{eq:lorenz}
        (x_k, \lc{\bmz}{\bmd}(k)) \coloneqq \left( \frac{1}{D}\sum_{i=1}^k d_{\pi(i)}, \sum_{i=1}^k z_{\pi(i)} \right).
    \end{equation}
    We define the set of points on this curve, $\lc{\bmz}{\bmd}(x),\ x \in [0,1]$, as the \emph{Lorenz curve} of vector $\bmz$ with respect to $\bmd$.
\end{definition}
Components $x_k$ are rescaled by $D$ so that comparison of curves with unequal dimensions is possible.
In fact, the Lorenz curves $\lc{\bmz}{\bmd}$ and $\lc{\bmz \otimes \bmd}{\bmd \otimes \bmd}$, where $\otimes$ denotes the Kronecker product, are identical.
Furthermore, a Lorenz curve $\lc{\bmz}{\bmd}(x)$ is always concave in $x$, since it consists of $d$ line segments each with slope $(\tilde{\bmz}^\downarrow)_i$ for $i=1,\dots,d$ which is non-increasing by definition.
Finally, the points on the interior of the Lorenz curve connecting line segments of different slopes are called \emph{elbows}.

A vector $\bmy$ is said to \emph{$\bmd$--majorize} another vector $\bmx$ if and only if the Lorenz curve $\lc{\bmy}{\bmd}$ lies above Lorenz curve $\lc{\bmx}{\bmd}$, thus reducing $\bmd$--majorization into a finite set of inequalities.
\begin{theorem}\label{thm:dmajor}
    Let $\bmx, \bmy, \bmd \in \reals^d$, such that the components of $\bmd$ are positive. 
    Then, $\bmx \prec_{\bmd} \bmy$ if and only if $\lc{\bmx}{\bmd}(x) \leq \lc{\bmy}{\bmd}(x)$ for all $x \in [0,1]$ with strict equality at $x=1$.
\end{theorem}
A restatement of the theorem including more equivalent conditions and a proof are provided in~\cref{app:major}.
An example of comparison between different Lorenz curves is illustrated in~\cref{fig:lctoy}.
\begin{figure}
    \centering
    \includegraphics[height=5cm]{figs/lctoy.pdf}
    \caption{Example of different Lorenz curves for quasi-probability vectors under $\bmd$-majorization.
    Vectors $\bmy$ and $\bmd$ are simply probability distributions.
    The curve corresponding to vector $\bmd$ is always the line segment connecting $(0,0)$ and $(1,1)$, so that any other Lorenz curve lies above it, for example $\bmx \prec_{\bmd} \bmd$.
    Curves $L_k(\bmx|\bmd)$ and $L_k(\bmy|\bmd)$ intersect, so $\bmx \nprec_{\bmd} \bmy$ as well as $\bmy \nprec_{\bmd} \bmx$.
    \nick{Recast in terms of magic - replace with~\cref{fig:test}?}
    }
    \label{fig:lctoy}
\end{figure}
